---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

# First we need to preprocess the data and set some variables like the name of the day of the week (which is different based on language, but used later on)


```{r}
## Step 0a: Define general variables like working directory. You can change this to something that suits you to test the script!
workingdir            <- "C:/Users/Jim Wittermans/Documents/GitHub/RepData_PeerAssessment1/activity/" ## Emptied for my own privacy ;-)
file_name             <- "activity.csv"
saturday              <- "zaterdag"         # Change this to your own language; warning: case sensitive!
sunday                <- "sunday"           # Change this to your own language; warning: case sensitive!

## Step 0b: Set global variables / load libraries
setwd(workingdir)
library(lattice)

######################################################################################### --- PART 1 --- #########################################################
## Step 1: Read and prepare the CSV as defined in step 1 
rep_dataset            <- read.csv(file_name, head=TRUE, colClasses=c("integer", "character", "integer"), na.strings="NA")
rep_dataset$date       <- as.Date(rep_dataset$date)
rep_dataset_prepared   <- subset(rep_dataset, !is.na(rep_dataset$steps))
```


## What is mean total number of steps taken per day?

```{r}
## Step 2: Get the sum per day, using tapply, but only with filled values.
steps_per_day          <- tapply(rep_dataset_prepared$steps, rep_dataset_prepared$date, sum, na.rm=TRUE, simplify=T)
steps_per_day          <- steps_per_day[!is.na(steps_per_day)]

## Step 3: Generate Histogram (Breaks = Length(steps_per_day) indicates how many different bars have to appear, based on the amount of different records)
hist(x=steps_per_day,
     ylab="Frequency",
     xlab="Total Steps per Day",
     main="Frequency of steps per day",     
     col="lightblue",
     border=0,
     breaks=length(steps_per_day))

## Step 4: Report the mean and median
mean(steps_per_day)
median(steps_per_day)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
